---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
date: "2023-03-20"
---


## Loading and preprocessing the data
### 1.1. Downloading and unzipping the necessary files
```{r, echo=TRUE}
URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if(!file.exists("activity.zip")){
    download.file(URL, destfile = "activity.zip")
}

if(file.exists("activity.zip")){
    unzip("activity.zip")
}
```
### 1.2. loading necessary packages
```{r, echo=TRUE, results='hide'}
library(dplyr)
library(lattice)
library(grDevices)
```
### 1.3. Reading the data
```{r, echo=TRUE}
activity <- read.csv("activity.csv", sep = ",")
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?

### 2. Histogram of the total number of steps taken each day
```{r, echo=TRUE}
with.na <-aggregate(steps ~ date, data= activity, sum, na.rm = TRUE)
hist(with.na$steps, xlab= "total number of steps",breaks = 20, main = "The total number of steps taken each day")
```
### 3. Mean and median number of steps taken each day
```{r, echo=TRUE}
N <- summary(with.na)
paste("With Na,", N[4,2], N[3,2])
```

## What is the average daily activity pattern?

### 4. Time series plot of the average number of steps taken
```{r, echo=TRUE}
av.steps <-aggregate(steps ~ interval, data= activity, mean, na.rm = TRUE)
with(av.steps, plot(interval, steps, type ="l", main="Time series plot of the average number of steps taken", col="red", lwd=2))
```
### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
max.interval <- av.steps[which.max(av.steps$steps),]
paste("The 5-minute interval that contains the maximum number of steps is", max.interval$interval)
```
## Imputing missing values

### strategy: replacing NA values with their mean values corresponding to the interval
```{r, echo=TRUE}
paste("the number of Na values in the original dataset is", sum(is.na(activity)))
na.steps <- activity[which(is.na(activity$steps)),]
na.steps$steps <- av.steps$steps
activity.new <- activity
activity.new[which(is.na(activity.new$steps)),] <- na.steps
```
### 6.1 the difference of mean and median values of total steps taken each day after imputing Na values
```{r, echo=TRUE}
activity.new.aggreagted <- aggregate(steps ~ date, data = activity.new, sum)
M <- summary(activity.new.aggreagted)
paste("without Na", M[4,2], M[3,2])
```
### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=TRUE}
hist(activity.new.aggreagted$steps, xlab = "Total number of steps", breaks = 20,main = "The total number of steps taken each day")
```

## Are there differences in activity patterns between weekdays and weekends?
### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo=TRUE}
activity.vnew <- activity.new
activity.vnew$weekdays <- weekdays(activity.new$date)
activity.vnew$weekdays[activity.vnew$weekdays%in%c("Saturday", "Sunday")] <- "weekend"
activity.vnew$weekdays[activity.vnew$weekdays%in%c("Monday", "Tuesday","Wednesday","Friday", "Thursday")] <- "weekday"
activity.aggregated <- aggregate(steps ~ interval + weekdays, data=activity.vnew, mean)
xyplot(activity.aggregated$steps~activity.aggregated$interval|activity.aggregated$weekdays, layout=c(1,2), type="l", main= "Average Number of Steps Taken (across all weekday days or weekend days)",xlab = "Interval", ylab = "steps")
```
